---
title: "Simulations"
author: "Hadrien Lorenzo"
date: "10/09/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ddsPLS2)
library(missMDA)
source("functions0.R")
source('~/Documents/GitHub/Koh_Lanta/kohLanta2_fun.R')
source('~/Documents/GitHub/Koh_Lanta/MI_PLS.R')

## Koh-Lanta
load("../data/Simus/KohLantaVSNipals/Design1/ddsPLS.RData")
# save(df,sel_Y,R_pred,rmseStdS,paras,file="../data/Simus/KohLantaVSNipals/Design1/ddsPLS.RData")
ddspls_1_df1 <- cbind("ddsPLS",df)
names(ddspls_1_df1) <- c("method",names(df))
load("../data/Simus/KohLantaVSNipals/Design1/ddsPLS2.RData")
# save(df,sel_Y,R_pred,rmseStdS,paras,file="../data/Simus/KohLantaVSNipals/Design1/ddsPLS2.RData")
ddspls_1_df2 <- cbind("ddsPLS",df)
names(ddspls_1_df2) <- c("method",names(df))
ddspls_1_df <- rbind(ddspls_1_df1)#,ddspls_1_df2)

## MULTIPLE IMP PLS
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/MI_NIPALS.RData")
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/MI_NIPALS2.RData")
load("../data/Simus/KohLantaVSNipals/Design1/MI_NIPALS.RData")
miPLS_1_df1 <- data.frame(cbind("MI-NIPALS",df))
names(miPLS_1_df1) <- c("method",names(df))
load("../data/Simus/KohLantaVSNipals/Design1/MI_NIPALS2.RData")
miPLS_1_df2 <- data.frame(cbind("MI-NIPALS",df))
names(miPLS_1_df2) <- c("method",names(df))
miPLS_1_df <-  rbind(miPLS_1_df1)#,miPLS_1_df2)


## SIMPLE IMP PLS
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/SIMPLE_NIPALS.RData")
load("../data/Simus/KohLantaVSNipals/Design1/SIMPLE_NIPALS.RData")
pls_1_df1 <- data.frame(cbind("NIPALS",df))
names(pls_1_df1) <- c("method",names(df))
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/SIMPLE_NIPALS2.RData")
load("../data/Simus/KohLantaVSNipals/Design1/SIMPLE_NIPALS2.RData")
pls_1_df2 <- data.frame(cbind("NIPALS",df))
names(pls_1_df2) <- c("method",names(df))
pls_1_df <- rbind(pls_1_df1)#,pls_1_df2)


## MEAN IMP PLS
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/MEAN_NIPALS.RData")
load("../data/Simus/KohLantaVSNipals/Design1/MEAN_NIPALS.RData")
meamPLS_1_df1 <- data.frame(cbind("MEAN-NIPALS",df))
names(meamPLS_1_df1) <- c("method",names(df))
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/MEAN_NIPALS2.RData")
load("../data/Simus/KohLantaVSNipals/Design1/MEAN_NIPALS2.RData")
meamPLS_1_df2 <- data.frame(cbind("MEAN-NIPALS",df))
names(meamPLS_1_df2) <- c("method",names(df))
meamPLS_1_df <- rbind(meamPLS_1_df1)#,meamPLS_1_df2)

all_1_df <- rbind(ddspls_1_df,miPLS_1_df,pls_1_df,meamPLS_1_df)
all_1_df[,2] <- as.numeric(all_1_df[,2])
all_1_df[,4] <- as.numeric(all_1_df[,4])
cols <- RColorBrewer::brewer.pal(8,"Set1")
cols[6] <- "brown"
angle <- c(-45,45,90,-35)
density <- c(30,35,40,45)

source('~/Documents/GitHub/Koh_Lanta/plot_ddspls_nipals_1.R')

```

## First design of simulations

The first design considers 
\begin{equation}
\begin{array}{c c c}
\mathbf{A} =\sqrt{1-\sigma^2}
\left(\begin{array}{ccc}
\alpha_3\boldsymbol{1}_{3,50} & \boldsymbol{0}_{3,50} &  \boldsymbol{0}_{3,900}\\
\boldsymbol{0}_{2,50} & \alpha_2\boldsymbol{1}_{2,50} & \boldsymbol{0}_{2,900}\\
\end{array} \right)_{(5,1000)}
&\text{ and }&
\mathbf{D}=\sqrt{1-\sigma^2}
\left(\begin{array}{ccc}
\alpha_3\boldsymbol{1}_{3,1} & \boldsymbol{0}_{3,1} & \boldsymbol{0}_{3,1} \\ 
\boldsymbol{0}_{2,1} & \alpha_2\boldsymbol{1}_{2,1} & \boldsymbol{0}_{2,1} \\
\end{array} \right)_{(5,3)},
\end{array}
\label{equ:design_1}
\end{equation}
where $\forall k \in \mathbb{N}^\star, \ \alpha_k=1/\sqrt{k}$, $\sqrt{1-\sigma^2}=0.99$ and 
$$\boldsymbol{\psi}=(\boldsymbol{\phi}',\boldsymbol{\epsilon}'_{1\dots100}/\sigma,
\boldsymbol{\epsilon}'_{101\dots1000},\boldsymbol{\xi}'_{1\dots2}/\sigma,\xi_{3})'\sim \mathcal{N}\left(\mathbf{0},\mathbb{I}_{5+1000+3}\right).$$
The corresponding latent variable model is then:
\begin{equation}
\begin{array}{ccc}
x_j=\left\{\begin{array}{rcl}
\sqrt{1-\sigma^2}(\phi_1+\phi_2+\phi_3)/\sqrt{3}+\epsilon_j&\text{ for } & j=1\ldots 50 \\
\sqrt{1-\sigma^2}(\phi_4+\phi_5)/\sqrt{2}+\epsilon_j&\text{ for } & j=51\ldots 100 \\ 
\epsilon_j&\text{ for } & j=101\ldots 1000 \\ 
\end{array}\right.
&\text{ and }&
\left\{\begin{array}{l}
y_1=\sqrt{1-\sigma^2}(\phi_1+\phi_2+\phi_3)/\sqrt{3}+\xi_1\\
y_2=\sqrt{1-\sigma^2}(\phi_4+\phi_5)/\sqrt{2}+\xi_2\\ 
y_3=\xi_3 
\end{array}\right.
\end{array}
\end{equation}

## MCAR

```{r,echo=F,eval=F}
n <- 100
sqrt_1_minus_sig2 <- 0.9

p1 <- 10
p <- 2000
q <- 3
n_test <- 1000
props <- c(0,0.1,0.25,0.4,0.6)
nNAs <- round(n*p*props)
N_simus <- 10
id_simus <- 1:N_simus
B <- 20
M <- 5
R_max <- 5
paras <- expand.grid(nNAs,id_simus,B,M,n,p,q,1,sqrt_1_minus_sig2,n_test)
colnames(paras) <- c("nNA","id","B","M","n","p","q","design","sqrt_1_minus_sig2","n_test")

set.seed(1)
data_test <- get_design_1_na(n_test,sqrt_1_minus_sig2,p,p1 = p1,q,0)
rmseStdS <- matrix(NA,nrow(paras),q)
R_pred <- NULL
sel_X <- NULL
sel_Y <- NULL
errMin <- 1e-7


ID_BATCH <- 1

dev.new()

ids <- as.numeric(rownames(ddspls_1_df[which(ddspls_1_df$nNA==120000),]))

for(i in ids){#1:nrow(paras)){
  cat(i);cat(" | ")
  pa <- paras[i,]
  nNA <- pa[1]
  set.seed(i+N_simus*(ID_BATCH-1))
  data <- get_design_1_na(n,sqrt_1_minus_sig2,p,p1,q,as.numeric(nNA))
  
  # Koh-Lanta
  res <- KohLanta(data$X,data$Y,B = B,M = M,verbose=T)
  est <- predict(res,data_test$X)
  R_pred[[i]] <- res$R$Prediction
  sel_X[[i]] <- unlist(lapply(res$models,function(ll){ll$Selection$X}))
  sel_Y[[i]] <- unlist(lapply(res$models,function(ll){ll$Selection$Y}))
  
  ## MI PLS
  # models <- MI_PLS(data$X,data$Y,R=R_max,M=M,proper = T,errMin = errMin)
  # est <- list(est=predict_MI_PLS(models,data_test$X))
  # R_pred[[i]] <- lapply(models,function(mm){mm$R})
  
  # ## I PLS
  # models <- MI_PLS(data$X,data$Y,R=R_max,M=1,proper=F,errMin=errMin)
  # est <- list(est=predict_MI_PLS(models,data_test$X))
  # R_pred[[i]] <- lapply(models,function(mm){mm$R})
  
  ## MEAN PLS
  # x_mean <- data$X
  # mus_x <- colMeans(x_mean,na.rm = T)
  # idNA <- which(is.na(data$X),arr.ind = T)
  # idNAii <- which(is.na(data$X),arr.ind = F)
  # x_mean[idNAii] <- mus_x[idNA[,2]]
  # models <- MI_PLS(x_mean,data$Y,R=R_max,M=1,proper = F,errMin = errMin)
  # est <- list(est=predict_MI_PLS(models,data_test$X))
  # R_pred[[i]] <- lapply(models,function(mm){mm$R})
  
  ## missMDA
  ncomp <- estim_ncpPCA(data$X, ncp.min = 0, ncp.max = 5,verbose = T)
  res.imp <- imputePCA(geno, ncp = 2, scale = TRUE,method = "Regularized", row.w = NULL, coeff.ridge = 1,threshold = 1e-06, seed = NULL, nb.init = 1, maxiter = 1000)
  
  rmseStdS[i,] <- sqrt(colMeans((est$est-data_test$Y)^2))/apply(data_test$Y,2,sd)
  
  df <- data.frame(list(nNA=rep(paras[,1],q),y_j=unlist(lapply(1:q,function(jj){rep(paste("y",jj),length(paras[,1]))})),rmseSTD=as.vector(rmseStdS)))
  # par(mfrow=c(1,3))
  boxplot(rmseSTD~nNA,df[which(df$y_j=="y 1"),])
  # boxplot(rmseSTD~nNA,df[which(df$y_j=="y 2"),])
  # boxplot(rmseSTD~nNA,df[which(df$y_j=="y 3"),])
}

```

```{r,echo=F,fig.width=5,fig.height=1.5}
postscript("~/Documents/GitHub/Koh_Lanta/nipasl_ddsPLS_1.eps",
           onefile=TRUE, horizontal=F,width=9, height=2.5,pointsize=3)
plot_1()
dev.off()

```

## MNAR

```{r,echo=F}
n <- 100
sqrt_1_minus_sig2 <- 0.99
p <- 300
q <- 3
n_test <- 1000
quantiles <- c(0.1,1,10)
d <- 100
xx <- rnorm(1e5)
ths <- c(-3.1434578060861,-2.35185906809072,-1.27897347797873)#sort(xx)[length(xx)*quantiles/d]
N_simus <- 20
id_simus <- 1:N_simus
B <- 20
M <- 5
R_max <- 5
paras <- expand.grid(ths,id_simus,B,M,n,p,q,1,sqrt_1_minus_sig2,n_test)
colnames(paras) <- c("th","id","B","M","n","p","q","design","sqrt_1_minus_sig2","n_test")
set.seed(1)
data_test <- get_design_1_na_MNAR(n_test,sqrt_1_minus_sig2,p,q,threshold = NA)
rmseStdS <- matrix(NA,nrow(paras),q)
R_pred <- NULL
sel_X <- NULL
sel_Y <- NULL


ID_BATCH <- 1


for(i in 1:nrow(paras)){
  cat(i);cat(" | ")
  pa <- paras[i,]
  th <- pa[1]
  set.seed(i+N_simus*(ID_BATCH-1))
  data <- get_design_1_na_MNAR(n,sqrt_1_minus_sig2,
                          p,q,threshold = th)
  ## Koh-Lanta
  # res <- KohLanta(data$X,data$Y,B = B,M = M,verbose=T)
  # est <- predict(res,data_test$X)
  # R_pred[[i]] <- res$R$Prediction
  # sel_X[[i]] <- unlist(lapply(res$models,function(ll){ll$Selection$X}))
  # sel_Y[[i]] <- unlist(lapply(res$models,function(ll){ll$Selection$Y}))
  
  ## MI PLS
  models <- MI_PLS(data$X,data$Y,R=R_max,M=M,proper = T,errMin = errMin)
  est <- list(est=predict_MI_PLS(models,data_test$X))
  R_pred[[i]] <- lapply(models,function(mm){mm$R})
    
  ## MEAN PLS
  # x_mean <- data$X
  # mus_x <- colMeans(x_mean,na.rm = T)
  # idNA <- which(is.na(data$X),arr.ind = T)
  # idNAii <- which(is.na(data$X),arr.ind = F)
  # x_mean[idNAii] <- mus_x[idNA[,2]]
  # models <- MI_PLS(x_mean,data$Y,R=R_max,M=1,proper = F,errMin = errMin)
  # est <- list(est=predict_MI_PLS(models,data_test$X))
  # R_pred[[i]] <- lapply(models,function(mm){mm$R})
  
  # ## I PLS
  # models <- MI_PLS(data$X,data$Y,R=R_max,M=1,proper=F,errMin=errMin)
  # est <- list(est=predict_MI_PLS(models,data_test$X))
  # R_pred[[i]] <- lapply(models,function(mm){mm$R})
  
  rmseStdS[i,] <- sqrt(colMeans((est$est-data_test$Y)^2))/apply(data_test$Y,2,sd)
  
  df <- data.frame(list(th=rep(paras[,1],q),y_j=unlist(lapply(1:q,function(jj){rep(paste("y",jj),length(paras[,1]))})),rmseSTD=as.vector(rmseStdS)))
  # par(mfrow=c(1,3))
  boxplot(rmseSTD~th,df[which(df$y_j=="y 1"),])
  # boxplot(rmseSTD~th,df[which(df$y_j=="y 2"),])
  # boxplot(rmseSTD~th,df[which(df$y_j=="y 3"),])
}

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ddsPLS2)
source("functions0.R")
source('~/Documents/GitHub/Koh_Lanta/kohLanta2_fun.R')
source('~/Documents/GitHub/Koh_Lanta/MI_PLS.R')

## Koh-Lanta
load("../data/Simus/KohLantaVSNipals/Design1/ddsPLS_MNAR.RData")
# save(df,sel_Y,R_pred,rmseStdS,paras,file="../data/Simus/KohLantaVSNipals/Design1/ddsPLS_MNAR.RData")
ddspls_1_df1 <- cbind("ddsPLS",df)
names(ddspls_1_df1) <- c("method",names(df))
# load("../data/Simus/KohLantaVSNipals/Design1/ddsPLS_MNAR2.RData")
# # save(df,sel_Y,R_pred,rmseStdS,paras,file="../data/Simus/KohLantaVSNipals/Design1/ddsPLS_MNAR2.RData")
# ddspls_1_df2 <- cbind("ddsPLS",df)
# names(ddspls_1_df2) <- c("method",names(df))
ddspls_1_df <- rbind(ddspls_1_df1)#,ddspls_1_df2)

## MULTIPLE IMP PLS
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/MI_NIPALS_MNAR.RData")
load("../data/Simus/KohLantaVSNipals/Design1/MI_NIPALS_MNAR.RData")
miPLS_1_df1 <- data.frame(cbind("MI-NIPALS",df))
names(miPLS_1_df1) <- c("method",names(df))
miPLS_1_df <-  rbind(miPLS_1_df1)

## MEAN IMP PLS
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/MEAN_NIPALS_MNAR.RData")
load("../data/Simus/KohLantaVSNipals/Design1/MEAN_NIPALS_MNAR.RData")
meamPLS_1_df1 <- data.frame(cbind("MEAN-NIPALS",df))
names(meamPLS_1_df1) <- c("method",names(df))
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/MEAN_NIPALS_MNAR2.RData")
# load("../data/Simus/KohLantaVSNipals/Design1/MEAN_NIPALS_MNAR2.RData")
# meamPLS_1_df2 <- data.frame(cbind("MEAN-NIPALS",df))
# names(meamPLS_1_df2) <- c("method",names(df))
meamPLS_1_df <- rbind(meamPLS_1_df1)#,meamPLS_1_df2)

## SIMPLE IMP PLS
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/SIMPLE_NIPALS_MNAR.RData")
load("../data/Simus/KohLantaVSNipals/Design1/SIMPLE_NIPALS_MNAR.RData")
pls_1_df1 <- data.frame(cbind("NIPALS",df))
names(pls_1_df1) <- c("method",names(df))
# save(df,R_pred,file="../data/Simus/KohLantaVSNipals/Design1/SIMPLE_NIPALS_MNAR2.RData")
# load("../data/Simus/KohLantaVSNipals/Design1/SIMPLE_NIPALS_MNAR2.RData")
# pls_1_df2 <- data.frame(cbind("NIPALS",df))
# names(pls_1_df2) <- c("method",names(df))
pls_1_df <- rbind(pls_1_df1)#,pls_1_df2)

all_1_df <- rbind(ddspls_1_df,miPLS_1_df,pls_1_df,meamPLS_1_df)
all_1_df[,2] <- as.numeric(all_1_df[,2])
all_1_df[,4] <- as.numeric(all_1_df[,4])
cols <- RColorBrewer::brewer.pal(8,"Set1")
cols[6] <- "brown"
angle <- c(-45,45,90,-35)
density <- c(30,35,40,45)

source('~/Documents/GitHub/Koh_Lanta/plot_ddspls_nipals_1.R')

```

```{r,echo=F,fig.width=5,fig.height=1.5}
postscript("~/Documents/GitHub/Koh_Lanta/nipasl_ddsPLS_1_MNAR.eps",
           onefile=TRUE, horizontal=F,width=9, height=2.5,pointsize=3)
plot_1_MNAR()
dev.off()

```
